### 2015-05-13 whoop.tex
update in summarisation

\medskip\noindent\textbf{Generating Loop and Procedure Summaries }
%
Early in the development of \whoop we experimented with reasoning about recursion-free drivers by resorting to full inlining.  We found that this did not scale to large drivers: when we tried to apply \whoop to the r8169 ethernet driver ($>$7000 lines of C code and use of recursion) with all functions fully inlined, we quickly exhausted the memory limits of our experimental platform.  \ADComment{Is it thus the case that r8169 did not exhibit any recursion?}  The C-to-Boogie translation produced, after inlining, hundreds of thousands of lines of Boogie IVL code, consuming a lot of memory and leading to gigantic verification conditions that the Z3 theorem prover could not handle effectively.  \ADComment{I rephrased the previous sentence, because if I remember correctly we did get to the point of invoking Z3.  But maybe I imagined that - perhaps we could not even generate the VC?}  Furthermore, some entry points invoke recursive helper functions, which cannot be fully inlined.

\ADComment{I think the following is the weakest part of this (in general very good) section at the moment.  My concerns are that (a) we have not really invented anything new, we just applied Houdini having guessed candidate preconditions, postconditions and loop invariants based on heuristics, and (b) we have not given much intuition for what these candidates look like.  Could you give an example candidate, using mathematical notation?  For instance, if entry point $T$ has a procedure $f$ that releases mutex $m$, then when instrumenting $T$ with symbol $i$ (using the notation from \S\ref{sec:symbolicpairwise}) we might guess $m \notin \currentlockset{i}$ as a postcondition for $\instrument{f}{i}$.  It would be ideal if you could be as specific as possible about the candidates we generate.}

To make our analysis scalable, and to cater for recursion, we developed \emph{watchdog summarization}, a novel summarization technique that allows us to analyze drivers in a scalable fashion and without destroying precision.
Watchdog summarization uses the Houdini~\cite{flanagan2001houdini} invariant inference algorithm (available inside Boogie) to automatically compute summaries (pre- and post-conditions and loop invariants) from a pool of \emph{candidate} invariants.  The Boogie verifier then automatically checks whether each procedure meets its summary, abstracting each called procedure with its summary in the process; Boogie also checks whether loops respect their invariants and uses the invariants to over-approximate the effects of loops.

The candidate invariants are based on the aforementioned global watchdog variables that \whoop instruments for watchdog race-checking. The key idea behind watchdog summarization is to automatically generate candidate invariants for each memory region in the granularity of watched accesses. This reduces the false positives from reporting errors if only e.g.\ a field in memory region $\mathit{MR}$ is racing, but the rest of the fields in $\mathit{MR}$ are not racing.

@@

\ADComment{In the above, can you make it super-clear that it is fine for the candidates to be wrong: Houdini will just kick them out?  You might want to look at the comment on this in the GPUVerify 2012 OOPSLA paper; I remember writing something nice along these lines, feel free to re-use it.  If you search for Houdini you'll find it.}

>>>>

\medskip\noindent\textbf{Generating Loop and Procedure Summaries }
%
Early in the development of \whoop we experimented with reasoning about recursion-free drivers by resorting to full inlining.  We found that this did not scale to large drivers: when we tried to apply \whoop to the r8169 ethernet driver (see \S\ref{evaluation}), we quickly exhausted the memory limits of our experimental platform as the driver exhibited recursion, which cannot be fully inlined and blows up.

\ADComment{We have not given much intuition for what these candidates look like.  Could you give an example candidate, using mathematical notation?  For instance, if entry point $T$ has a procedure $f$ that releases mutex $m$, then when instrumenting $T$ with symbol $i$ (using the notation from \S\ref{sec:symbolicpairwise}) we might guess $m \notin \currentlockset{i}$ as a postcondition for $\instrument{f}{i}$.  It would be ideal if you could be as specific as possible about the candidates we generate.}

To make our analysis scalable without sacrificing precision, and to cater for recursion, we use the Houdini~\cite{flanagan2001houdini} invariant inference algorithm to automatically compute summaries (pre- and post-conditions and loop invariants) from a pool of \emph{candidate} invariants. The Boogie verifier then automatically checks whether each procedure meets its summary, abstracting each called procedure with its summary in the process. Boogie also checks whether loops respect their invariants and uses the invariants to over-approximate the effects of loops.

The candidate invariants are based on the aforementioned watchdog constants that \whoop instruments for watchdog race-checking. Our candidate generation rules are purely heuristic. The key idea is to automatically generate candidates for each memory region in the granularity of watched accesses. This reduces the false positives from reporting errors if only e.g.\ a field in memory region $\mathit{MR}$ is racing, but the rest of the fields in $\mathit{MR}$ are not racing.

@@

We emphasize that the candidate invariants generated by \whoop are just that: candidates. The tool is free to speculatively generate candidates that later turn out to be incorrect: these are simply discarded by Houdini. A consequence is that incorrect or unintended candidates generated due to bugs in \whoop cannot compromise the soundness of verification.
